DOM Review Summary

### 2024-03-15 Analysis/ReviewSummary/ReviewSummary.qmd

review_Objectives |>
  filter(Status == "Complete") |> 
  select(ID, Obj_Methods, Obj_Future, Obj_Spatial, Obj_Hypothesis) |>
  pivot_longer(cols = c(Obj_Methods, Obj_Future, 
                        Obj_Spatial, Obj_Hypothesis), 
               names_to = "Objective", 
               values_to = "Stated") |>
  mutate(Objective = str_extract(Objective, "(?<=Obj\\_).+"),
         Stated = case_when(Stated == "Y" ~ TRUE,
                            Stated == "N" ~ FALSE)) |>
  left_join(select(studyDetails_clean, ID, Year, YearStrata), by = "ID")

@@

@@

>>>>

objectives_Clean <- review_Objectives |>
  filter(Status == "Complete") |> 
  select(ID, Obj_Methods, Obj_Future, Obj_Spatial, Obj_Hypothesis) |>
  rowwise() |>
  mutate(Obj_None = case_when(all(c(Obj_Methods, Obj_Future,
                                    Obj_Spatial, Obj_Hypothesis) == "N") ~ "Y",
                              any(c(Obj_Methods, Obj_Future,
                                    Obj_Spatial, Obj_Hypothesis) == "Y") ~ "N")) |>
  pivot_longer(cols = c(Obj_Methods, Obj_Future, 
                        Obj_Spatial, Obj_Hypothesis, Obj_None), 
               names_to = "Objective", 
               values_to = "Stated") |>
  mutate(Objective = str_extract(Objective, "(?<=Obj\\_).+"),
         Stated = case_when(Stated == "Y" ~ TRUE,
                            Stated == "N" ~ FALSE)) |>
  left_join(select(studyDetails_clean, ID, 
                   Year, YearStrata), 
            by = "ID")
```

```{r Objectives_frequency}
#| echo: false

objectives_frequency <- objectives_Clean |>
  mutate(Total = length(unique(ID))) |>
  filter(Stated == TRUE) |>
  summarise(Articles = n(),
            .by = c(Objective, Total)) |>
  mutate(Proportion = Articles/Total) |> 
  arrange(-Proportion)
```

The most common objective was **Testing Hypotheses**, with `{r} max(objectives_frequency$Proportion)`% of articles examining pre-determined relationships. Prediction, whether spatial or temporal, was rare.

```{r Objectives_frequencyPlot}
#| echo: false

objectives_frequency |>
  mutate(Objective = fct(Objective,
                         levels = c("Hypothesis", "Methods",
                                    "Future", "Spatial",
                                    "None"))) |>

  ggplot() +
  geom_col(aes(x = Objective, y = Articles, fill = Objective)) +
  scale_fill_manual(values = c("darkorchid", "deepskyblue3",
                               "coral2", "goldenrod", "gray60")) +
  labs(title = "Number of articles by objective",
       subtitle = "'None' indicates that none of the objectives were stated,\ngenerally this indicates a focus on occupancy dynamics parameters",
       x = "Stated objective", y = "Number of articles") +
  theme(legend.position = "none")

@@

Little variation was shown between the objectives of papers by different authorship categories.

```{r Objective_frequencyAuthorship}
#| echo: false

objectives_frequencyAuthorship <- objectives_Clean |>
  left_join(select(studyDetails_clean, ID, AuthorCategories), by = "ID") |>
  separate_rows(AuthorCategories, sep = ", ") |>
  filter(Stated == TRUE) |>
  mutate(Total = length(unique(ID)), .by = AuthorCategories) |>
  summarise(Articles = n(),
            .by = c(Objective, AuthorCategories, Total)) |>
  mutate(Proportion = Articles/Total) |>
  complete(Objective, AuthorCategories, 
           fill = list(Articles = 0,
                       Proportion = 0))

# Plot frequency faceted by authorship
objectives_frequencyAuthorship |>
  mutate(Objective = fct(Objective,
                         levels = c("Hypothesis", "Methods",
                                    "Future", "Spatial",
                                    "None"))) |>

  ggplot() +
  geom_col(aes(x = Objective, y = Proportion, fill = Objective)) +
  scale_fill_manual(values = c("darkorchid", "deepskyblue3",
                               "coral2", "goldenrod", "gray60")) +
  facet_wrap(facets = vars(AuthorCategories)) +
  labs(title = "Proportion of articles with each objective",
       subtitle = "'None' indicates that none of the listed objectives were stated,\ngenerally this indicates a focus on simply estimating occupancy dynamics",
       x = "Stated objective", y = "Number of articles") +
  theme(legend.position = "bottom")
```

@@

Objective categories have changed in frequency over time, with more variation in objectives in recent years relative to the early days of DOMs

```{r Objectives_byYear}
#| echo: false

objectives_Clean |>
  filter(!is.na(YearStrata)) |>
  filter(Stated == TRUE) |>
  mutate(Total = length(unique(ID)), .by = YearStrata) |>
  summarise(Articles = n(), .by = c(Objective, YearStrata, Total)) |>
  mutate(Proportion = Articles/Total) |>
  complete(Objective, YearStrata, fill = list(Articles = 0, Proportion = 0)) |>
  mutate(Objective = fct(Objective,
                         levels = c("Hypothesis", "Methods",
                                    "Future", "Spatial",
                                    "None"))) |>

  ggplot() +
  geom_col(aes(x = YearStrata, y = Proportion, fill = Objective),
           position = "dodge") +
  scale_fill_manual(values = c("darkorchid", "deepskyblue3",
                               "coral2", "goldenrod", "gray60")) +
  labs(title = "Objective categories over time",
       subtitle = "'None' indicates that none of the listed objectives were stated,\ngenerally this indicates a focus on simply estimating occupancy dynamics",
       y = "Proportion of articles", x = "Year Strata")
